CDSR Human Data/Modelling Paper

### 2013-09-12 code/scenarios/cdsr/trunk/papers/ACS2013/main.tex

Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.  To identify these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects. Research question 2 address the contextual nature of CDSRs. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

To examine these questions, we developed two experimental tasks:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were asked to rate how well the marked location fitted with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

@@

\subsection{Trial Type 1 - Design}
In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.  Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.

\subsection{Trial Type 1 - Results}

One well known metric for analysising interannotator agreement is the \textbf{Cohen's kappa coefficient}, $\kappa$. Cohen's kappa measures the agreement between two annotators who each classify a number of items into two or more mutually exclusive classes. $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement. In order to use Cohen's kappa to measure the agreement between subject responses in our type 1 trials we generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used. Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e. there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

@@

\begin{figure}
%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
\centerline{
\begin{tabular}{ccc}
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
\end{tabular}
}
\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
\label{fig:type-1-hists}
\end{figure}

The analysis of the type 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.

\subsection{Trial Type 2 - Design} 
In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.  The location of these points was informed by the results of a pilot study. In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region. In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}). One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped. In the classroom domain the expected sweetspot varied between the points depending on the stimuli. In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets. P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.

\subsection{Trial Type 2 - Results}

@@

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.

>>>>

Consider the following regions (shown in Figure \ref{fig:examples}): the front of a classroom, safety in a military engagement, or the geographical feature of a bottleneck.  To determine these regions, one must identify the physical boundaries (e.g., walls, bodies of water), the objects within the environment, their orientation and their functional use (e.g., desks oriented toward a whiteboard), and other agents' intentions and capabilities (e.g., enemy tanks can move and have a range of attack).  Furthermore, these regions exist at different levels of spatial resolution.  Consider a neighborhood in a city, whose boundaries expand and contract over time with changes to the surrounding residents and businesses.

@@

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining variation in CDSR placement within different situtations. In a situation exhibiting clear contextual clues (e.g. determining a safe zone in a battle situation where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR. In a less well-defined context, one might expect little to no agreement between human communicators.  Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

%Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextually dependent then one would expect a variation %across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a %high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement %between subjects. Research question 2 address the contextual nature of CDSRs. Research question 2 addresses the contextual nature of CDSRs %in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

To examine these questions, we developed two experimental tasks:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the marked location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

% kate: vocabulary clarification: situation vs. context??

@@

\subsection{Task 1 - Design}
In the first set of trials, the subject was asked to draw a polygon that defined the extent of the CDSR given or to select a no region option.  The drawing was done on the diagram in all stimuli with the classroom photographs provided for reference in those trials.  Subjects drew polygons by clicking to position the vertices with lines being automatically generated between each point indicated as the subject drew.  There was no upper bound placed on the number of line segments drawn but subjects were required to provide at least three line segments or to choose no region before continuing to the next stimulus.

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
% but want to talk to John about it during the telecon tomorrow first

We use \textbf{Cohen's kappa coefficient}, $\kappa$, to measure the agreement between two regions as follows.  We generated a set of 47,000 test points that were uniformly distributed across each stimuli and used each subject's polygon to classify the test points as being inside or outside the polygon. After classifying the test points for each subject we then computed an average pairwise kappa score for each subject with all the other subjects on a per stimuli basis. In this calculation if either of the subjects in a pair entered a no region response for a particular stimuli a $\kappa$ value of 0 was used.  $\kappa$ ranges from 0 indicating no agreement beyond chance to 1 indicating complete agreement.

% Klenk: This is where I left off.  I think we can drop the histogram and talk.

Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e. there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

@@

%Kate: this is a place (above) that I think we could combine/cut.  I think we can talk about the non-typical classrooms more concisely.  Let's
% say something like "some of the layouts were non-typical -> less experience with the context -> we expect lower agreement -> this is what 
% we see" then we can say that if we remove these examples, agreement on the more typical scence is good.  So suggestions are
% (1) condense discussion (2) make it clear exactly what we expect for the non typical scenes and WHY this is tied back to our claim
% about the role of context in CDSRs.  I am happy to make these changes if John agrees that I am not horribly misinterpreting his fine results

The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.

%\begin{figure}
%%\includegraphics[width=0.4\textwidth]{figures/turrets-plus-one.jpeg}
%\centerline{
%\begin{tabular}{ccc}
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist.pdf}\caption{}\label{fig:subj-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/stimuli-kappa-hist.pdf}\caption{}\label{fig:stim-kappa}\end{subfigure}&
%\begin{subfigure}[b]{0.3\textwidth}\includegraphics[width=\textwidth]{figures/subj-kappa-hist-protostim.pdf}\caption{}\label{fig:subj-kappa-proto}\end{subfigure}
%\end{tabular}
%}
%\caption{(a) Histogram of average pairwise subject kappa score across all stimuli by subject, (b) histogram of average pairwise subject kappa scores for each stimuli, (c) histogram of average pairwise subject kappa score excluding non-prototypical stimuli by subject.}
%\label{fig:type-1-hists}
%\end{figure}

\begin{table}
\centerline{
\begin{tabular}{c c}
Expected Relationship & Human Results  \\
\hline
\hline
area(g) > area(h) & 291.2 > 276.2 (p<.003) \\
maximumX(f) > maximumX(e) & 176.1 > 167.1   (p<.47) \\
minimumX(e-flipped) > minimumX(f-flipped) & 336.1 > 321.3 (p<.12) \\
maximumY(i) > maxmimumY(j) & 304.64 > 265.04 (p<.03) \\
maximumY(j) > maxmimumY(k) & 265.04 > 251.71 (p<.47) \\
maximumY(k) = maxmimumY(l) & 265.04 != 290.03 (p<.04) \\
\hline
\end{tabular}
}
\label{tab:region-location-tests}
\caption{Comparison of the expected relationships to the mean human responses.}
\end{table}

\subsection{Task 2 - Design} 
In the second set of trials, a location was marked on the visual stimuli (using a red dot) and the subject was asked to rank on a five point likert scale their judgement of how well the location matched the described spatial region.  The location of these points was informed by the results of a pilot study. In the final study they were positioned such that one of the domain's points, called the \textbf{sweetspot}, was located near the expected prototypical center of the described spatial region. In the classroom domain the three points were located along the horizontal line that bisected the stimuli with P1 positioned on the left of the stimuli, P2 in the center and P3 on the right of the stimuli (as shown in Figure \ref{fig:classroom-points}). One of the aims of this configuration of points was to provide comparable data points between the trials where the classroom stimuli were horizontally flipped. In the classroom domain the expected sweetspot varied between the points depending on the stimuli. In the map domain P1 was positioned in the top left corner of the stimuli,  P2 was the expected sweetspot and was located within the expected bottleneck region for both stimuli and P3 was located in the bottom right corner of the stimuli (see Figure \ref{fig:bottleneck-points}).  In the military engagement domain the three test points were located on the vertical line that bisected the stimuli: P1 was located near the top of the image in the proximity of the enemy tanks, P2 was located near the center of the image such that in some stimuli it was in a no-man's land position and in other stimuli (specifically stimuli (i) in the friendly tanks have moved forward and (l) where there is an extra friendly turret) it could be considered to be in a protected area, and P3 was located near the bottom of the stimuli and consequently was always located behind the friendly tanks/turrets. P3 was the expected sweetspot in the military engagement domain but it was also hoped that P2 rating would differ between the stimuli where it was in no-mans land and where it recieved some protection.

%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though

\subsection{Task 2 - Results}

@@

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli. %lending support for question 2 ...
